---
title: "Machine Learning Project"
author: "Hugo Cam"
date: "Tuesday, December 16, 2014"
output: html_document
---

## Introduction

The goal is to predict the manner in which persons did the exercise. 
This is the "classe" variable in the training set. I will use any of the other variables to predict with.
We will also use the prediction model to predict 20 different test cases.

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

Based on the data recorded from several people using Activity Monitors, I will try to construct a macine learning algorithm to predict future outcomes on the test data.
I am using Random Forest Modeling, and got over 99% accuracy on my training set of 70% of the total data. 


## Loading and cleaning Data

After downloaded the data files, I will review the data and cleanning it.

```{r}
library(caret)
testing <- read.csv("pml-testing.csv", na.strings = c("NA", ""))
training <- read.csv("pml-training.csv", na.strings = c("NA", ""))
dim(testing); dim(training)
head(training)
```


Removing columns X, username, time stamps fields and new_window field. Columns 1 to 6.

```{r}
removeColumns <- as.integer(c(1, 2, 3, 4, 5, 6))
training <- training[, -removeColumns]
testing <- testing[, -removeColumns]

```

Eliminating columns with missing values
```{r}
numberNAs <- apply(training, 2, function(x) { sum(is.na(x)) })
training2 <- training[, which(numberNAs == 0)]
dim(training2)
```

## Building data sets for training and testing data sets.

Using 70% for training and 30% for testing.  None generated for testing since that set is already provided.

```{r}
inTrain <- createDataPartition(y = training2$classe, p = 0.7, list = FALSE)
trainingSet <- training2[inTrain, ]
testingSet  <- training2[-inTrain, ]
```

## Fitting the model and Predict

Training, fit the model, outcome is Classe, predicted with all remaining variables
method rain forest, the data is trainingSet  

Parameter tuning, cross-validation and simple bootstrap resampling is used for line 3 in the algorithm above. The function trainControl can be used to specifiy the type of resampling: cross-validation

```{r}
trControl = trainControl(method = "cv", number = 4)
## fit the model
## install.packages("e1071")
modelFit <- train(trainingSet$classe ~ ., data = trainingSet, method = "rf", trControl = trControl)
modelFit
```

Calculation the errors using the testing Set. It is the number of valid prediction divide by total testing set.

```{r}
predicted <- predict(modelFit, testingSet)
SampleError <- sum(predicted == testingSet$classe)/nrow(testingSet)
SampleError
```

Generating data for the prediction vector for the Assigment Submission
```{r}
answers <- predict(modelFit, testing)
answers
```

I will use the following function to create the files required to submit the answers

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)
```